---
title: "baseline paper"
author: "Ale Garcia"
date: "2025-03-26"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
library(foreign)

theme_set(theme_minimal())
```

## WAVE 1 - COHORT 1

379 Rows. 

Before 06/04/25, 9 rows with no identifiers. I asked JAA to re_uploaded the data with all identifiers, and he did on 06/05/25. Today is 06/17/25 and I confirmed that all rows have identifiers. Although I am habing issues using import and here. Using foreign does work. 

### Loading the dataset (cohort 1 - wave 1)
```{r}
raw_par_w1c1 <- import(here("nopublish", "caregivers_assessment_1_co1.sav"), setclass = "tbl_df")

par_w1c1 <- raw_par_w1c1 %>% 
  select(-2:-4) %>% # selecting out columns with non-relevant data
  clean_names() %>% # function that formats variables' names
  arrange(record_id) %>% # ordering participants ids in descending order
  mutate(cohort = 1) %>% # creating cohort variable, in case 
  select(1, cohort, everything()) # moving cohort var to place 2

data.frame(table(par_w1c1$record_id)) # no duplicated ids

sum(is.na(par_w1c1$record_id)) # all rows have an identifier 

view_df(par_w1c1) # this shows values and value labels and prefer not to respond is 99, no aplica 77
```

# WAVE 1 - COHORT 2

846 Rows. 2 rows with no identifiers. 

### Loading the dataset (cohort 2 - wave 1)
```{r}
raw_par_w1c2 <- import(here("nopublish", "caregivers_assessment_1_co2.sav"), setclass = "tbl_df")

par_w1c2 <- raw_par_w1c2 %>% 
  select(-2:-4) %>% # selecting out columns with non-relevant data
  clean_names() %>% # function that formats variables' names
  arrange(record_id) %>% # ordering participants ids in descending order
  mutate(cohort = 2) %>% # creating cohort variable, in case 
  select(1, cohort, everything())

data.frame(table(par_w1c2$record_id)) # no duplicated ids

sum(is.na(par_w1c2$record_id)) # all rows have an identifier 

view_df(par_w1c2) # this shows values and value labels and prefer not to respond is 99, no aplica 77
```

## Merging parent cohort 1 and 2 (this is only wave 1)

1225 rows, 11 with no identifiers.  

Two ids for a 2701
2701.1
2701.2

```{r}
par_w1 <- full_join(par_w1c1, par_w1c2) %>% 
  arrange(record_id)

data.frame(table(par_w1$record_id)) # no duplicated ids

sum(is.na(par_w1$record_id)) # all rows have an identifier 

view_df(par_w1) # this shows values and value labels and prefer not to respond is 99, no aplica 77
```

```{r}
par_w1 %>%
  haven::write_sav(here("data", "caregivers_co1_co2_w1.sav"))
```


0202_1
0202_2

#######################################################################################
STUDENTS
#######################################################################################

## WAVE 1 - COHORT 1

373 Rows. 2 rows with no identifiers. 

### Loading the student dataset (cohort 1 - wave 1)
```{r}
raw_stu_w1c1 <- import(here("nopublish", "01. SPSS Students Assessment 1 Cohort 1.ESP.sav"), setclass = "tbl_df")

stu_w1c1 <- raw_stu_w1c1 %>% 
  select(-2:-4) %>% # selecting out columns with non-relevant data
  clean_names() %>% # function that formats variables' names
  arrange(record_id) %>% # ordering participants ids in descending order
  mutate(cohort = 1) %>% # creating cohort variable, in case 
  select(1, cohort, everything()) # moving cohort var to place 2

data.frame(table(stu_w1c1$record_id)) # no duplicated ids

sum(is.na(stu_w1c1$record_id)) # 2 rows don't have an identifier 
```

## WAVE 1 - COHORT 2

840 Rows. All rows with identifiers. 

### Loading the student dataset (cohort 2 - wave 1)
```{r}
raw_stu_w1c2 <- import(here("nopublish", "01. Students Assessment 1 Cohorte 2.sav"), setclass = "tbl_df")

stu_w1c2 <- raw_stu_w1c2 %>% 
  select(-2:-4) %>% # selecting out columns with non-relevant data
  clean_names() %>% # function that formats variables' names
  arrange(record_id) %>% # ordering participants ids in descending order
  mutate(cohort = 2, # creating cohort variable, in case 
         survey_other = as.character(survey_other)) %>% # can't join without this change
  select(1, cohort, everything()) # moving cohort var to place 2

data.frame(table(stu_w1c2$record_id)) # no duplicated ids

sum(is.na(stu_w1c2$record_id)) # all rows with identifier 
```

## Merging student cohort 1 and 2 (this is only wave 1)

1213 rows, 2 with no identifiers.  

```{r}
stu_w1 <- full_join(stu_w1c1, stu_w1c2) %>% 
  arrange(record_id)

data.frame(table(stu_w1$record_id)) # no duplicated ids

sum(is.na(stu_w1$record_id)) # 2 rows don't have an identifier 
```

```{r}
view_df(stu_w1)
```

### 06.04.25: with JAA identified that:

1038_a = 9 años, niño en cuarto grado cambiar a 1038.1
1038_b = 11 años, niña en quinto grado cambiar a 1038.1

```{r}
which(is.na(stu_w1$record_id)) # gives the rows
  
stu_w1$record_id[1212] <- "1038.1"
stu_w1$record_id[1213] <- "1038.2"

stu_w1 <- stu_w1 %>%
  arrange(record_id)
```


#######################################################################################
EDUCATORS
#######################################################################################

## WAVE 1 - COHORT 1

107 Rows. All rows with identifiers. 

### Loading the educator dataset (cohort 1 - wave 1)
```{r}
raw_edu_w1c1 <- import(here("nopublish", "01. SPSS Educators Assessment 1 Cohort 1.ESP.sav"), setclass = "tbl_df")

edu_w1c1 <- raw_edu_w1c1 %>% 
  select(-2:-4) %>% # selecting out columns with non-relevant data
  clean_names() %>% # function that formats variables' names
  arrange(record_id) %>% # ordering participants ids in descending order
  mutate(cohort = 1) %>% # creating cohort variable, in case 
  select(1, cohort, everything()) # moving cohort var to place 2

data.frame(table(edu_w1c1$record_id)) # no duplicated ids

sum(is.na(edu_w1c1$record_id)) # all rows with an identifier 
```

## WAVE 1 - COHORT 2

268 Rows. All rows with identifiers. 

### Loading the student dataset (cohort 2 - wave 1)
```{r}
raw_edu_w1c2 <- import(here("nopublish", "01. SPSS Educators Assessment 1 Cohort 2.ESP.sav"), setclass = "tbl_df")

edu_w1c2 <- raw_edu_w1c2 %>% 
  select(-2:-4) %>% # selecting out columns with non-relevant data
  clean_names() %>% # function that formats variables' names
  arrange(record_id) %>% # ordering participants ids in descending order
  mutate(cohort = 2) %>% # creating cohort variable, in case 
  select(1, cohort, everything()) # moving cohort var to place 2

data.frame(table(edu_w1c2$record_id)) # no duplicated ids

sum(is.na(edu_w1c2$record_id)) # all rows with an identifier  
```

## Merging educator cohort 1 and 2 (this is only wave 1)

375 rows, all with  identifiers.  

```{r}
edu_w1 <- full_join(edu_w1c1, edu_w1c2) %>% 
  arrange(record_id)

data.frame(table(edu_w1$record_id)) # no duplicated ids

sum(is.na(edu_w1$record_id)) # all rows with an identifier 
```

```{r}
view_df(edu_w1)
```


#######################################################################################
PRINCIPALS
#######################################################################################

## WAVE 1 - COHORT 1

11 Rows. All rows with identifiers. 

### Loading the principal dataset (cohort 1 - wave 1)
```{r}
raw_pri_w1c1 <- import(here("nopublish", "01. SPSS Principals Assessment 1 Cohort 1.ESP.sav"), setclass = "tbl_df")

pri_w1c1 <- raw_pri_w1c1 %>% 
  select(-2:-4) %>% # selecting out columns with non-relevant data
  clean_names() %>% # function that formats variables' names
  arrange(record_id) %>% # ordering participants ids in descending order
  mutate(cohort = 1) %>% # creating cohort variable, in case 
  select(1, cohort, everything()) # moving cohort var to place 2

data.frame(table(pri_w1c1$record_id)) # no duplicated ids

sum(is.na(pri_w1c1$record_id)) # all rows with an identifier 
```

## WAVE 1 - COHORT 2

19 Rows. All rows with identifiers. 

### Loading the student dataset (cohort 2 - wave 1)
```{r}
raw_pri_w1c2 <- import(here("nopublish", "01. SPSS Principals Assessment 1 Cohort 2.ESP.sav"), setclass = "tbl_df")

pri_w1c2 <- raw_pri_w1c2 %>% 
  select(-2:-4) %>% # selecting out columns with non-relevant data
  clean_names() %>% # function that formats variables' names
  arrange(record_id) %>% # ordering participants ids in descending order
  mutate(cohort = 2) %>% # creating cohort variable, in case 
  select(1, cohort, everything()) # moving cohort var to place 2

data.frame(table(pri_w1c2$record_id)) # no duplicated ids

sum(is.na(pri_w1c2$record_id)) # all rows with an identifier  
```

## Merging principal cohort 1 and 2 (this is only wave 1)

30 rows. All rows with identifiers. 

```{r}
pri_w1 <- full_join(pri_w1c1, pri_w1c2) %>% 
  arrange(record_id)

data.frame(table(pri_w1$record_id)) # no duplicated ids

sum(is.na(pri_w1$record_id)) # all rows with an identifier  
```

```{r}
view_df(pri_w1)
```

# Full datasets for all participant-types

### parents

female caregivers = 1057
male caregivers = 163
99 = 4 
NA = 1 
total = 1225

intro001_1:
1, Soy su madre = 920
2, Soy su padre = 131
3, Otra relación = 169

demo001_1:
1, Femenino
2, Masculino
99, Prefiero no responder	

Sleuthing:
99 = co1: 835 (tia), co2: 2122 (dad), 2542 (?), 2824(mom) = 2 female, 1 male, 1 ?
NA = co1: 438 (mom) = 1 female

Conclusions of my sleuthing: 
female caregivers = 1060
male caregivers = 164
1 row with 99, id 2542 (?)
total = 1225

```{r}
par_w1
view_df(par_w1)

par_w1 %>%
  	count(demo001_1)

# to identify who are the 99
par_w1 %>%
  	filter(demo001_1 == 99)

# to identify who is the na
par_w1 %>% filter(is.na(demo001_1))

par_w1 %>% filter(record_id == "835") 
par_w1 %>% filter(record_id == "2122") 
par_w1 %>% filter(record_id == "2542") # lots of prefer not to respond
par_w1 %>% filter(record_id == "2824") 
par_w1 %>% filter(record_id == "438") # lots of prefer not to respond

```
### students

girls = 602
boys = 611

total = 1213

```{r}
stu_w1
view_df(stu_w1)

stu_w1 %>%
  	count(demo002)
```

```{r}
stu_w1 %>% 
  mean(demo001, na.rm = TRUE)

# age
mean(stu_w1$demo001, na.rm = TRUE)
sd(stu_w1$demo001, na.rm = TRUE)

# gender
stu_w1 %>%
  count(demo002) %>%
  mutate(percent = n / sum(n) * 100)

# grade
stu_w1 %>%
  count(demo003) %>%
  mutate(percent = n / sum(n) * 100)

# # of people they live with
mean(stu_w1$demo004, na.rm = TRUE)
sd(stu_w1$demo004, na.rm = TRUE)

stu_w1 %>%
  filter(demo004 >= 7) %>%
  nrow()

stu_w1 %>%
  filter(demo006input_a >= 10) %>%
  nrow()

# job
stu_w1 %>%
  count(demo006) %>%
  mutate(percent = n / sum(n) * 100)

stu_w1 %>%
  count(school001) %>%
  mutate(percent = n / sum(n) * 100)


```
```{r}

```



### teachers

demo001_3
0, Femenino
1, Masculino
99, Prefiero no responder

Female = 317
male = 55
99 = 2 (co1: 1174, co2: 2180) --> they self-report as primary caregivers

eyeballing: 
demo001_3 = 2 rows with 99 and 1 row with na

```{r}
edu_w1
view_df(edu_w1)

edu_w1 %>%
  	count(demo001_3)

# to identify who are the 99
edu_w1 %>% filter(demo001_3 == "99") 

# to identify who is the na
edu_w1 %>% filter(is.na(demo001_3))

edu_w1 %>% filter(record_id == "1376") # no data (only 0 and na)
edu_w1 %>% filter(record_id == "1174")
edu_w1 %>% filter(record_id == "2180")
```


### principals 
demo001_4
1, Femenino
2, Masculino
99, Prefiero no responder

1, female = 19
2, male = 11

```{r}
pri_w1
view_df(pri_w1)

pri_w1 %>%
  	count(demo001_4)

# to identify who are the 99
pri_w1 %>% filter(demo001_4 == "99") 

# to identify who is the na
pri_w1 %>% filter(is.na(demo001_4)) # no na in this dataset
```




