---
title: "baseline paper"
author: "Ale Garcia"
date: "2025-03-26"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)

theme_set(theme_minimal())
```

## WAVE 1 - COHORT 1

379 Rows. 9 rows with no identifiers. 

### Loading the dataset (cohort 1 - wave 1)
```{r}
raw_par_w1c1 <- import(here("nopublish", "01. SPSS Caregivers Assessment 1 Cohort 1.ESP.sav"), setclass = "tbl_df")

par_w1c1 <- raw_par_w1c1 %>% 
  select(-2:-4) %>% # selecting out columns with non-relevant data
  clean_names() %>% # function that formats variables' names
  arrange(record_id) %>% # ordering participants ids in descending order
  mutate(cohort = 1) %>% # creating cohort variable, in case 
  select(1, cohort, everything()) # moving cohort var to place 2

data.frame(table(par_w1c1$record_id)) # no duplicated ids

sum(is.na(par_w1c1$record_id)) # 9 rows don't have an identifier 
```

# WAVE 1 - COHORT 2

846 Rows. 2 rows with no identifiers. 

### Loading the dataset (cohort 1 - wave 1)
```{r}
raw_par_w1c2 <- import(here("nopublish", "01. SPSS Caregivers Assessment 1 Cohort 2.ESP.sav"), setclass = "tbl_df")

par_w1c2 <- raw_par_w1c2 %>% 
  select(-2:-4) %>% # selecting out columns with non-relevant data
  clean_names() %>% # function that formats variables' names
  arrange(record_id) %>% # ordering participants ids in descending order
  mutate(cohort = 2) %>% # creating cohort variable, in case 
  select(1, cohort, everything())

data.frame(table(par_w1c2$record_id)) # no duplicated ids

sum(is.na(par_w1c2$record_id)) # 2 rows don't have an identifier 
```

## Merging parent cohort 1 and 2 (this is only wave 1)

1225 rows, 11 with no identifiers.  

Two ids for a 2701
2701.1
2701.2

```{r}
par_w1 <- full_join(par_w1c1, par_w1c2) %>% 
  arrange(record_id)

data.frame(table(par_w1$record_id)) # no duplicated ids

sum(is.na(par_w1$record_id)) # 11 rows don't have an identifier 
```

```{r}
view_df(par_w1)
```

#######################################################################################
STUDENTS
#######################################################################################

## WAVE 1 - COHORT 1

373 Rows. 2 rows with no identifiers. 

### Loading the student dataset (cohort 1 - wave 1)
```{r}
raw_stu_w1c1 <- import(here("nopublish", "01. SPSS Students Assessment 1 Cohort 1.ESP.sav"), setclass = "tbl_df")

stu_w1c1 <- raw_stu_w1c1 %>% 
  select(-2:-4) %>% # selecting out columns with non-relevant data
  clean_names() %>% # function that formats variables' names
  arrange(record_id) %>% # ordering participants ids in descending order
  mutate(cohort = 1) %>% # creating cohort variable, in case 
  select(1, cohort, everything()) # moving cohort var to place 2

data.frame(table(stu_w1c1$record_id)) # no duplicated ids

sum(is.na(stu_w1c1$record_id)) # 2 rows don't have an identifier 
```

## WAVE 1 - COHORT 2

840 Rows. All rows with identifiers. 

### Loading the student dataset (cohort 2 - wave 1)
```{r}
raw_stu_w1c2 <- import(here("nopublish", "01. Students Assessment 1 Cohorte 2.sav"), setclass = "tbl_df")

stu_w1c2 <- raw_stu_w1c2 %>% 
  select(-2:-4) %>% # selecting out columns with non-relevant data
  clean_names() %>% # function that formats variables' names
  arrange(record_id) %>% # ordering participants ids in descending order
  mutate(cohort = 2, # creating cohort variable, in case 
         survey_other = as.character(survey_other)) %>% # can't join without this change
  select(1, cohort, everything()) # moving cohort var to place 2

data.frame(table(stu_w1c2$record_id)) # no duplicated ids

sum(is.na(stu_w1c2$record_id)) # all rows with identifier 
```

## Merging student cohort 1 and 2 (this is only wave 1)

1213 rows, 2 with no identifiers.  

```{r}
stu_w1 <- full_join(stu_w1c1, stu_w1c2) %>% 
  arrange(record_id)

data.frame(table(stu_w1$record_id)) # no duplicated ids

sum(is.na(stu_w1$record_id)) # 2 rows don't have an identifier 
```

```{r}
view_df(stu_w1)
```

#######################################################################################
EDUCATORS
#######################################################################################

## WAVE 1 - COHORT 1

107 Rows. All rows with identifiers. 

### Loading the educator dataset (cohort 1 - wave 1)
```{r}
raw_edu_w1c1 <- import(here("nopublish", "01. SPSS Educators Assessment 1 Cohort 1.ESP.sav"), setclass = "tbl_df")

edu_w1c1 <- raw_edu_w1c1 %>% 
  select(-2:-4) %>% # selecting out columns with non-relevant data
  clean_names() %>% # function that formats variables' names
  arrange(record_id) %>% # ordering participants ids in descending order
  mutate(cohort = 1) %>% # creating cohort variable, in case 
  select(1, cohort, everything()) # moving cohort var to place 2

data.frame(table(edu_w1c1$record_id)) # no duplicated ids

sum(is.na(edu_w1c1$record_id)) # all rows with an identifier 
```

## WAVE 1 - COHORT 2

268 Rows. All rows with identifiers. 

### Loading the student dataset (cohort 2 - wave 1)
```{r}
raw_edu_w1c2 <- import(here("nopublish", "01. SPSS Educators Assessment 1 Cohort 2.ESP.sav"), setclass = "tbl_df")

edu_w1c2 <- raw_edu_w1c2 %>% 
  select(-2:-4) %>% # selecting out columns with non-relevant data
  clean_names() %>% # function that formats variables' names
  arrange(record_id) %>% # ordering participants ids in descending order
  mutate(cohort = 2) %>% # creating cohort variable, in case 
  select(1, cohort, everything()) # moving cohort var to place 2

data.frame(table(edu_w1c2$record_id)) # no duplicated ids

sum(is.na(edu_w1c2$record_id)) # all rows with an identifier  
```

## Merging educator cohort 1 and 2 (this is only wave 1)

375 rows, all with  identifiers.  

```{r}
edu_w1 <- full_join(edu_w1c1, edu_w1c2) %>% 
  arrange(record_id)

data.frame(table(edu_w1$record_id)) # no duplicated ids

sum(is.na(edu_w1$record_id)) # all rows with an identifier 
```

```{r}
view_df(edu_w1)
```


#######################################################################################
PRINCIPALS
#######################################################################################

## WAVE 1 - COHORT 1

11 Rows. All rows with identifiers. 

### Loading the principal dataset (cohort 1 - wave 1)
```{r}
raw_pri_w1c1 <- import(here("nopublish", "01. SPSS Principals Assessment 1 Cohort 1.ESP.sav"), setclass = "tbl_df")

pri_w1c1 <- raw_pri_w1c1 %>% 
  select(-2:-4) %>% # selecting out columns with non-relevant data
  clean_names() %>% # function that formats variables' names
  arrange(record_id) %>% # ordering participants ids in descending order
  mutate(cohort = 1) %>% # creating cohort variable, in case 
  select(1, cohort, everything()) # moving cohort var to place 2

data.frame(table(pri_w1c1$record_id)) # no duplicated ids

sum(is.na(pri_w1c1$record_id)) # all rows with an identifier 
```

## WAVE 1 - COHORT 2

19 Rows. All rows with identifiers. 

### Loading the student dataset (cohort 2 - wave 1)
```{r}
raw_pri_w1c2 <- import(here("nopublish", "01. SPSS Principals Assessment 1 Cohort 2.ESP.sav"), setclass = "tbl_df")

pri_w1c2 <- raw_pri_w1c2 %>% 
  select(-2:-4) %>% # selecting out columns with non-relevant data
  clean_names() %>% # function that formats variables' names
  arrange(record_id) %>% # ordering participants ids in descending order
  mutate(cohort = 2) %>% # creating cohort variable, in case 
  select(1, cohort, everything()) # moving cohort var to place 2

data.frame(table(pri_w1c2$record_id)) # no duplicated ids

sum(is.na(pri_w1c2$record_id)) # all rows with an identifier  
```

## Merging principal cohort 1 and 2 (this is only wave 1)

30 rows. All rows with identifiers. 

```{r}
pri_w1 <- full_join(pri_w1c1, pri_w1c2) %>% 
  arrange(record_id)

data.frame(table(pri_w1$record_id)) # no duplicated ids

sum(is.na(pri_w1$record_id)) # all rows with an identifier  
```

```{r}
view_df(pri_w1)
```






