---
title: "cfa paper"
author: "Alejandra Garcia Isaza"
date: "2025-07-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)
library(lavaan)

theme_set(theme_minimal())
```


### Loading the dataset (cohort 1 and 2 - wave 1)
```{r}
d_stu <- import(here("data", "students_co1_co2_w1.sav"), setclass = "tbl_df")

view_df(d_stu) # Not showing each value label separate. Values appear as range, prob due to haven stripping labels when I exported. 
```


### recoding missing variables as N/A
```{r include=FALSE}

# vector with missing values in dataset
missing_vals <- c(77, 99)

# function that returns true if values in vector are equal to missing_vals. The function takes a vector x, and specified values of missing data
recode_missing <- function(x, missing_vals = c(77, 99)) {
  test <- x %in% missing_vals
  ifelse(test, NA, x)
}

# function that recodes missing values to NA. The function takes a dataframe with variables with missing data, and specified values of missing data
recode_missing_df <- function(df, missing_vals = c(77, 99)) {
  modify(df, ~recode_missing(.x, missing_vals)) # here uses the function created above
}

d1_stu <- recode_missing_df(d_stu) # the function strips out variable labels
```

### Data prep: reverse scoring negatively worded items
SDQ values = 1, 2, 3
```{r}
d2_stu <- d1_stu %>%
  mutate(sdq007 = likert_reverse(sdq007, top = 3, bottom = 1),
         sdq011 = likert_reverse(sdq011, top = 3, bottom = 1),
         sdq014 = likert_reverse(sdq014, top = 3, bottom = 1),
         sdq021 = likert_reverse(sdq021, top = 3, bottom = 1),
         sdq025 = likert_reverse(sdq025, top = 3, bottom = 1))
```


### checking reverse scoring
```{r}
# d1_stu$sdq007
# d2_stu$sdq007
# Conclusion: it worked
```

### SDQ scales
emotional difficulties: sdq003, sdq008, sdq013, sdq016, sdq024 -> higher, worse emotional problems
conduct difficulties: sdq005, sdq007 (R), sdq012, sdq018, sdq022 -> higher, worse conduct problems
Hyperactivity scale: sdq002, sdq010, sdq015, sdq021 (R), sdq025 (R) -> higher, worse hyperactivity
Peer difficulties scale: sdq006, sdq011 (R), sdq014 (R), sdq019, sdq023 -> higher, worse peer diff
Prosocial scale: sdq001, sdq004, sdq009, sdq017, sdq020 -> higher, better prosociality

```{r}
emo_diff <- d2_stu %>%
  select(sdq003, sdq008, sdq013, sdq016, sdq024) 
alpha(emo_diff) # 0.62

cond_diff <- d2_stu %>%
  select(sdq005, sdq007, sdq012, sdq018, sdq022) 
alpha(cond_diff) # 0.49

hyper_diff <- d2_stu %>%
  select(sdq002, sdq010, sdq015, sdq021, sdq025) 
alpha(hyper_diff) # 0.57

peer_diff <- d2_stu %>%
  select(sdq006, sdq011, sdq014, sdq019, sdq023) 
alpha(peer_diff) # 0.35

prosocial <- d2_stu %>%
  select(sdq001, sdq004, sdq009, sdq017, sdq020) 
alpha(prosocial) # 0.52


tot_diff <- d2_stu %>%
  select(sdq003, sdq008, sdq013, sdq016, sdq024, sdq005, sdq007, sdq012, sdq018, sdq022, sdq002, sdq010, sdq015, sdq021, sdq025, sdq006, sdq011, sdq014, sdq019, sdq023) 
alpha(tot_diff) # 0.73
```


####sdq 1####
```{r}
sdq_1 <-("
#factor structure 
emo=~ sdq003 + sdq008 + sdq013 + sdq016 + sdq024
         cond=~ sdq005 + sdq007 + sdq012 + sdq018 + sdq022
         hyper=~ sdq002 + sdq010 + sdq015 + sdq021 + sdq025
         peer=~ sdq006 + sdq011 + sdq014 + sdq019 + sdq023
         pros=~ sdq001 + sdq004 + sdq009 + sdq017 + sdq020
         
         #factor covariances
         emo~~ cond + hyper + peer + pros
         cond ~~ hyper + peer + pros
         hyper ~~ peer + pros
         peer ~~ pros")
```


```{r}
sdq_1cfa<-cfa(sdq_1, data=d2_stu, estimator="DWLS", ordered = T)
summary(sdq_1cfa, fit.measures=T, standardized=T, rsquare=T)
```

ajuste = matriz var cov observada concuerda con the implied matrix chi2 

  P-value (Chi-square)                           0.000 terrible 

  Comparative Fit Index (CFI)                    0.902 trabajable
  Tucker-Lewis Index (TLI)                       0.890 trabajable

  RMSEA                                          0.046 trabajable
  
  SRMR                                           0.071 trabajable



latwnt variables = cargas factoriales 

peer =~                                                               
    sdq006            1.000                               0.346    0.346
    sdq011            0.736    0.108    6.826    0.000    0.254    0.254
    sdq014            0.901    0.095    9.471    0.000    0.311    0.311
    sdq019            1.727    0.145   11.917    0.000    0.597    0.597
    sdq023            0.960    0.096    9.992    0.000    0.332    0.332


but not too bad

cov y corr
cond ~~                                                               
    hyper              0.848 potencialment multicolineal, pero nada grave (>.95 seria porb)


Variances: p significativo == we are ok

R-sqrd == these are too low
sdq011            0.065
sdq014        0.097

el problema fundamental es el ajuste.


```{r}
qchisq(0.05, 265, lower.tail=F) # VALOR CRITICO AL TENER 256 GRADOS DE LIBERTAD, SALE EN EL OUTPUT. tENEMOS 934.880 IDEAL SERIA 303 PARA ABAJO
```


####sdq 2####
```{r}
sdq_2 <-("
#factor structure 
emo=~ sdq003 + sdq008 + sdq013 + sdq016 + sdq024
         cond=~ sdq005 + sdq007 + sdq012 + sdq018 + sdq022
         hyper=~ sdq002 + sdq010 + sdq015 + sdq021 + sdq025
         peer=~ sdq006 + sdq011 + sdq014 + sdq019 + sdq023
         pros=~ sdq001 + sdq004 + sdq009 + sdq017 + sdq020
         
         #factor covariances
         emo~~ cond + hyper + peer + pros
         cond ~~ hyper + peer + pros
         hyper ~~ peer + pros
         peer ~~ pros

         # Mod indices
         sdq011 ~~ sdq014
         sdq002 ~~ sdq010") # tienen algo que el modelo no puede capturar, estamos agregando una cov a los errores de estos items
```


```{r}
modindices(sdq_1cfa, sort. = T)
```

chi didn't imprve as much, SRMR tampoco,  pero CFI AND TLI, yes. 
```{r}
sdq_2cfa<-cfa(sdq_2, data=d2_stu, estimator="DWLS", ordered = T)
summary(sdq_2cfa, fit.measures=T, standardized=T, rsquare=T)
anova(sdq_1cfa, sdq_2cfa) # significant improvement in the fit indexes
```


1. Goodman = ya
2. 5 factors con 2 lat segundo nivel
3. Honduras paper
4. factores de seg = agarrar emo items + peer items sin segundo orden 
5. 1 solo factor 


####sdq 2####
```{r}
sdq_2 <-("
#factor structure 
emo=~ sdq003 + sdq008 + sdq013 + sdq016 + sdq024
         cond=~ sdq005 + sdq007 + sdq012 + sdq018 + sdq022
         hyper=~ sdq002 + sdq010 + sdq015 + sdq021 + sdq025
         peer=~ sdq006 + sdq011 + sdq014 + sdq019 + sdq023
         pros=~ sdq001 + sdq004 + sdq009 + sdq017 + sdq020
         int =~ peer + emo
         ext =~ cond + hyp
         
         #factor covariances
         emo~~ cond + hyper + peer + pros
         cond ~~ hyper + peer + pros
         hyper ~~ peer + pros
         peer ~~ pros") 
```