---
title: "Educators baseline & scales"
author: "Alejandra Garcia Isaza"
date: "2025-07-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)

theme_set(theme_minimal())
```


### Loading the dataset (cohort 1 and 2 - wave 1)
```{r}
d_edu <- import(here("data", "educators_co1_co2_w1.sav"), setclass = "tbl_df") # 375 teachers

view_df(d_edu) # Not showing each value label separate. Values appear as range, prob due to haven stripping labels when I exported. 
```


### recoding missing variables as N/A
```{r include=FALSE}

# vector with missing values in dataset
missing_vals <- c(77, 88, 99)

# function that returns true if values in vector are equal to missing_vals. The function takes a vector x, and specified values of missing data
recode_missing <- function(x, missing_vals = c(77, 88, 99)) {
  test <- x %in% missing_vals
  ifelse(test, NA, x)
}

# function that recodes missing values to NA. The function takes a dataframe with variables with missing data, and specified values of missing data
recode_missing_df <- function(df, missing_vals = c(77, 88, 99)) {
  modify(df, ~recode_missing(.x, missing_vals)) # here uses the function created above
}

d1_edu <- recode_missing_df(d_edu) # the function strips out variable labels
```


### Demo vars
```{r}
# age
mean(d1_edu$demo002_3, na.rm = TRUE)
sd(d1_edu$demo002_3, na.rm = TRUE)

# gender
d1_edu %>%
  	count(demo001_3) %>%
    mutate(percent = n / sum(n) * 100)

# to identify who is the na
d1_edu %>% filter(is.na(demo001_3))

# Sleuthing:
d1_edu %>% filter(record_id == "1174") # 
d1_edu %>% filter(record_id == "1376") # 
d1_edu %>% filter(record_id == "2180") # 

# Ed level
d1_edu %>%
  count(experien001_3) %>%
  mutate(percent = n / sum(n) * 100)

# years teaching
d1_edu %>%
  count(experien002_3) %>%
  mutate(percent = n / sum(n) * 100)

# how many schools you work in
d1_edu %>%
  count(intro002_3) %>%
  mutate(percent = n / sum(n) * 100)

# number of students
d1_edu %>%
  filter(studinfo001_3 == "980")
# ids 833, 1793, 2185, 2484 have a value of 0, which I am excluding
# id 2087	has a value of 980, which I am excluding
# id 2284 has a value of 1202, which I am excluding
# many other ids have very large numbers > 80

mean(d1_edu$studinfo001_3[!d1_edu$studinfo001_3 %in% c(0, 980, 1202)], na.rm = TRUE)
sd(d1_edu$studinfo001_3[!d1_edu$studinfo001_3 %in% c(0, 980, 1202)], na.rm = TRUE)
range(d1_edu$studinfo001_3[!d1_edu$studinfo001_3 %in% c(0, 52, 54, 58, 89, 98, 100, 102, 112, 150, 199, 200, 250, 300, 320, 380, 400, 420, 980, 1202)], na.rm = TRUE) # range 1-45

range(d1_edu$studinfo001_3[d1_edu$cohort == 1], na.rm = TRUE)
range(d1_edu$studinfo001_3[d1_edu$cohort == 2], na.rm = TRUE)

# income enough?
d1_edu %>%
  count(demo011_3) %>%
  mutate(percent = n / sum(n) * 100)

# food insecurity
d1_edu %>%
  count(demo012_3) %>%
  mutate(percent = n / sum(n) * 100)

# parientes now abroad
d1_edu %>%
  count(demo007_3) %>%
  mutate(percent = n / sum(n) * 100) # ~23%

# estado civil
d1_edu %>%
  count(demo003_3) %>%
  mutate(percent = n / sum(n) * 100)

# teachers with 0 children and descriptives of # of children
d1_edu %>% filter(demo004_3 == "0") # ~ 15%
range(d1_edu$demo004_3[!d1_edu$demo004_3 %in% c(31)], na.rm = TRUE)
mean(d1_edu$demo004_3[!d1_edu$demo004_3 %in% c(31)], na.rm = TRUE)
sd(d1_edu$demo004_3[!d1_edu$demo004_3 %in% c(31)], na.rm = TRUE)


# primary caregiver
d1_edu %>%
  count(demo008_3) %>%
  mutate(percent = n / sum(n) * 100)

# hours per day w/child on a weekday
d1_edu %>%
  count(demo009_3) %>%
  mutate(percent = n / sum(n) * 100)
```

### Issues
I identified very high numbers in the studinfo00x_3 -- some educators reported total students they had, not where they were homeroom teachers.
```{r}
range(d1_edu$studinfo001_3[d1_edu$cohort == "1"], na.rm = TRUE)
range(d1_edu$studinfo001_3[d1_edu$cohort == "2"], na.rm = TRUE)

d1_edu %>%
  filter(cohort == "1") %>%
  summarise(
    min_value = min(studinfo001_3, na.rm = TRUE),
    max_value = max(studinfo001_3, na.rm = TRUE)
  )

d1_edu %>%
  filter(cohort == "2") %>%
  summarise(
    min_value = min(studinfo001_3, na.rm = TRUE),
    max_value = max(studinfo001_3, na.rm = TRUE)
  )

d1_edu %>%
  filter(studinfo001_3 > 40) 

# no aplica
d1_edu %>%
  filter(intro001_3_77 == 1) 

# prefer not to respond
d1_edu %>%
  filter(intro001_3_99 == 1) 
```

### teacher Version of the Teacher–Student Relationship Inventory (T-TSRI; Ang, 2005):
Instrument has 3 factors: satisfaction, help, and conflict. We didn't include the conflict scale. 
```{r}
s_satrel <- d1_edu %>%
  select(tsri007_3, tsri008_3, tsri009_3, tsri010_3, tsri011_3) 
alpha(s_satrel) # 0.86
```


### code for creating scales
```{r}
teacher_scales <- d1_edu %>%
  rowwise() %>%
  mutate(s_satrel = mean(c(tsri007_3, tsri008_3, tsri009_3, tsri010_3, tsri011_3), na.rm = TRUE))
```


### Average scores gy gender
0 = woman, 1 = man
```{r}
# satisfaction with rel with students
teacher_scales %>%
  group_by(demo001_3) %>%
  summarise(
    mean_t_satrel = mean(s_satrel, na.rm = TRUE),
    sd_t_satrel = sd(s_satrel, na.rm = TRUE),
    n = n()
  )
```

### TSRI
No sig. difference between satirsfaction in relation based on teacher's gender. 
```{r}
# Checking model assumptions (ignoring clustering, though)
table(is.na(teacher_scales$s_satrel), useNA = "always") # 39 N/A

table(teacher_scales$demo001_3) # Very unequal group sizes (0 = 317, 1 = 55)

var.test(s_satrel ~ factor(demo001_3), teacher_scales) # variances are equal for externalizing

# Two-sided, automatic variance check
t.test(s_satrel ~ factor(demo001_3), teacher_scales)
```
# tsri008_3
Describiría mis relaciones con estos alumnos como positivas.

0 = almost never true; 2 = sometimes is true,  4 = almost always true

almost never/rarely = 7.74%
sometimes = 13%
frequently/almost always = 64.4
n/a = 15
```{r}
# primary caregiver
d1_edu %>%
  count(tsri008_3) %>%
  mutate(percent = n / sum(n) * 100)
```
