---
title: "initial cleaning"
output: html_document
date: "2023-11-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)

theme_set(theme_minimal())
```

# Loading the dataset (cohort 1 - wave 1)
```{r}
d_raw_w1 <- import(here("nopublish", "Cuidadores Medición 1 Cohorte 1.sav"), setclass = "tbl_df")
```

## WAVE 1

#initial cleaning (wave 1)
```{r}
d_w1 <- d_raw_w1 %>% 
  select(-4:-6) %>% # selecting out columns with non-relevant data
  clean_names() %>% # function that formats variables' names
  arrange(record_id) # ordering participants ids in descending order

# d_raw_w1 <- d_raw_c1[-c(1:5), ] # to remove rows of data
```

# Checking duplicated ids
```{r}
# data.frame(table(d_w1$record_id))
```

Duplicate ids:
1038 (confirmed in redcap: one caregiver completed survey for two children boy and girl)
in wave 2 participant only completed survey for 1 child (the boy)

# before recoding N/A
```{r}
view_df(d_w1)
```

I identified the following variables to convert to N/A:
77 = No aplica
98 = no estoy sugura
99 = prefer not to respond

# recoding missing variables as N/A
```{r include=FALSE}

# vector with missing values in dataset
missing_vals <- c(77, 98, 99)

# function that returns true if values in vector are equal to missing_vals. The function takes a vector x, and specified values of missing data
recode_missing <- function(x, missing_vals = c(77, 98, 99)) {
  test <- x %in% missing_vals
  ifelse(test, NA, x)
}

# function that recodes missing values to NA. The function takes a dataframe with variables with missing data, and specified values of missing data
recode_missing_df <- function(df, missing_vals = c(77, 98, 99)) {
  modify(df, ~recode_missing(.x, missing_vals)) # here uses the function created above
}

d_w1_clean <- recode_missing_df(d_w1) # the function strips out variable labels
```

# after recoding N/A
```{r}
# view_df(d_w1_clean)
```

##  SCALES 
_1 = wave 1
* = need reverse scoring

# SDQ
internalizing_1 = sdq003_1, sdq008_1, sdq013_1, sdq016_1, sdq024_1
externalizing_1 = sdq005_1, sdq007_1*, sdq012_1, 	sdq018_1, sdq022_1
hyperactivity_1 = sdq002_1, sdq010_1, sdq015_1, sdq021_1* , sdq025_1*
socialization_1 = sdq006_1, sdq011_1* , sdq014_1* , sdq019_1, sdq023_1
prosocial_1 = sdq001_1, sdq004_1, sdq009_1, sdq017_1, sdq020_1

# PAFAS 
(higher scores - lower construct) consistency_1 = pafas001_1, pafas003_1* , pafas004_1, pafas011_1*, 	pafas012_1
(higher scores - higher construct) coercion_1 = pafas005_1, pafas007_1, pafas009_1, pafas010_1, pafas013_1
(higher scores - higher construct) encouragement_1 = pafas002_1, pafas006_1, pafas008_1
(higher scores - higher construct) relquality_1 = pafas014_1, pafas015_1, pafas016_1, pafas017_1, pafas018_1
(higher scores - lower construct) famrels_1 = pafas019_1* , pafas020_1* , pafas021_1, pafas022_1
(higher scores - higher construct) teamwork_1 = pafas023_1 , pafas024_1* , pafas025_1

# Data prep: reverse scoring negatively worded items
```{r}
d_w1_clean2 <- d_w1_clean %>%
  mutate(sdq007_1 = likert_reverse(sdq007_1, top = 2, bottom = 0),
         sdq021_1 = likert_reverse(sdq021_1, top = 2, bottom = 0),
         sdq025_1 = likert_reverse(sdq025_1, top = 2, bottom = 0),
         sdq011_1 = likert_reverse(sdq011_1, top = 2, bottom = 0),
         sdq014_1 = likert_reverse(sdq014_1, top = 2, bottom = 0),
         pafas003_1 = likert_reverse(pafas003_1, top = 3, bottom = 0),
         pafas011_1 = likert_reverse(pafas011_1, top = 3, bottom = 0),
         pafas019_1 = likert_reverse(pafas019_1, top = 3, bottom = 0),
         pafas020_1 = likert_reverse(pafas020_1, top = 3, bottom = 0),
         pafas024_1 = likert_reverse(pafas024_1, top = 3, bottom = 0),
         )
```

# scale creation
```{r}
w1_scales <- d_w1_clean2 %>%
  rowwise() %>% 
  mutate(internalizing_1 = mean(c(sdq003_1, sdq008_1, sdq013_1, sdq016_1, sdq024_1), na.rm = TRUE),
         externalizing_1 = mean(c(sdq005_1, sdq007_1, sdq012_1, sdq018_1, sdq022_1), na.rm = TRUE),
         hyperactivity_1 = mean(c(sdq002_1, sdq010_1, sdq015_1, sdq021_1, sdq025_1), na.rm = TRUE),
         socialization_1 = mean(c(sdq006_1, sdq011_1, sdq014_1, sdq019_1, sdq023_1), na.rm = TRUE),
         prosocial_1 = mean(c(sdq001_1, sdq004_1, sdq009_1, sdq017_1, sdq020_1), na.rm = TRUE),
         consistency_1 = mean(c(pafas001_1, pafas003_1, pafas004_1, pafas011_1, pafas012_1), na.rm = TRUE),
         coercion_1 = mean(c(pafas005_1, pafas007_1, pafas009_1, pafas010_1, pafas013_1), na.rm = TRUE),
         encouragement_1 = mean(c(pafas002_1, pafas006_1, pafas008_1), na.rm = TRUE),
         relquality_1 = mean(c(pafas014_1, pafas015_1, pafas016_1, pafas017_1, pafas018_1), na.rm = TRUE),
         famrels_1 = mean(c(pafas019_1, pafas020_1, pafas021_1, pafas022_1), na.rm = TRUE),
         teamwork_1 = mean(c(pafas023_1, pafas024_1, pafas025_1), na.rm = TRUE),
         )

# view_df(w1_scales)
```
# checking reverse scoring
```{r}
# d_w1_clean$sdq007_1
# d_w1_clean$pafas003_1

# w1_scales$sdq007_1
# w1_scales$pafas003_1

# Conclusion: the reverse scoring worked. 
```


# internal consistency (alpha scores)
```{r}
# SDQ SCALES #
w1_scales %>%
  select(sdq003_1, sdq008_1, sdq013_1, sdq016_1, sdq024_1) %>%
  alpha() # internalizing 0.70

w1_scales %>%
  select(sdq005_1, sdq007_1, sdq012_1, sdq018_1, sdq022_1) %>%
  alpha() # externalizing 0.58

w1_scales %>%
  select(sdq002_1, sdq010_1, sdq015_1, sdq021_1, sdq025_1) %>%
  alpha() # hyperactivity 0.68

w1_scales %>%
  select(sdq006_1, sdq011_1, sdq014_1, sdq019_1, sdq023_1) %>%
  alpha() # socialization 0.35

w1_scales %>%
  select(sdq001_1, sdq004_1, sdq009_1, sdq017_1, sdq020_1) %>%
  alpha() # prosocial 0.60

# PAFAS SCALE #

w1_scales %>%
  select(pafas001_1, pafas003_1, pafas004_1, pafas011_1, pafas012_1) %>%
  alpha() # consistency_1 0.23 - warning: items 3 and 11 neg corr with total scale - why if it's already recoded?

# d_w1_clean %>%
#   select(pafas001_1, pafas003_1, pafas004_1, pafas011_1, pafas012_1) %>%
#   alpha() # consistency without recoding .39

w1_scales %>%
  select(pafas005_1, pafas007_1, pafas009_1, pafas010_1, pafas013_1) %>%
  alpha() # coercion_1 0.73

w1_scales %>%
  select(pafas002_1, pafas006_1, pafas008_1) %>%
  alpha() # encouragement_1 0.56

w1_scales %>%
  select(pafas014_1, pafas015_1, pafas016_1, pafas017_1, pafas018_1) %>%
  alpha() # relquality_1 0.89

w1_scales %>%
  select(pafas019_1, pafas020_1, pafas021_1, pafas022_1) %>%
  alpha() # famrels_1 0.63

w1_scales %>%
  select(pafas023_1, pafas024_1, pafas025_1) %>%
  alpha() # teamwork_1 0.54

##### these not added to w1_scales yet ######

# CORE MDM SCALES #
d_w1_clean %>%
	select(coremdm014_1:coremdm024_1) %>% # 12 items
  alpha() # 0.83

d_w1_clean %>%
	select(coremdm016_1, coremdm017_1, coremdm018_1, coremdm019_1) %>% # 4 items
  alpha() # 0.80 taps on problem solving/emotion regulation

d_w1_clean %>%
	select(coremdm013_1, coremdm014_1, coremdm015_1, coremdm020_1, coremdm021_1, coremdm022_1, coremdm023_1, coremdm024_1) %>% # 8 items
  alpha() # 0.71 taps on supervision as in noticing child's emotions, behaviors, whereabouts

# NEG problem solving #
d_w1_clean %>%
	select(esp008_1, esp009_1, esp010_1) %>% 
  alpha() # 0.59

# Family Communication Scale (FCS) scales
d_w1_clean %>%
	select(comfam001_1, comfam002_1, comfam003_1, comfam004_1, comfam005_1, comfam006_1) %>% 
  alpha() # 0.90 family communication scale (unidimensional)

d_w1_clean %>%
	select(comfam001_1, comfam002_1, comfam005_1) %>% 
  alpha() # 0.84 general comunication (chilean validation)

d_w1_clean %>%
	select(comfam003_1, comfam004_1, comfam006_1) %>% 
  alpha() # 0.84 emotional expression (chilean validation)

# parent involvement
d_w1_clean %>%
	select(involveduc001_1, involveduc002_1, involveduc003_1, involveduc004_1, involveduc005_1) %>% 
  alpha() # 0.76 plug-in involvement (with or without 5)

# parent-teacher rel
d_w1_clean %>%
	select(relteach001_1, relteach002_1, relteach003_1, relteach005_1, relteach006_1) %>% 
  alpha() # relation with teacher 0.86

# parent's school endorsement
d_w1_clean %>%
	select(endorschool001_1, endorschool002_1, endorschool003_1, endorschool004_1) %>% 
  alpha() # school endorsement 0.89

# Parent to parent relationships
d_w1_clean %>%
	select(relpar001_1, relpar002_1, relpar003_1, relpar004_1, relpar005_1, relpar006_1) %>% 
  alpha() # relation with other parents at school 0.87

```




############################################################################################################################################################################################################### WAVE 2 ################################################################# ############################################################################################################################################


# Loading the dataset (cohort 1 - wave 2)
```{r}
d_raw_w2 <- import(here("nopublish", "Cuidadores Medición 2 Cohorte 1.sav"), setclass = "tbl_df")
```

#initial cleaning (wave 2)
```{r}
d_w2 <- d_raw_w2 %>% 
  select(-4:-6) %>% # selecting out columns with non-relevant data
  clean_names() %>% # function that formats variables' names
  arrange(record_id) # ordering participants ids in descending order

# d_raw_w1 <- d_raw_c1[-c(1:5), ] # to remove rows of data
```

# Checking duplicated ids
```{r}
# data.frame(table(d_w2$record_id))
```
Duplicate ids:
no duplicated id, but oddly named ones:

202.1 (complete w1, complete w2)
202.2 (missing w1, complete w2)

228.2 (missing w1, complete w2) - 228_1 complete w1, missing w2 
403.2 (missing w1, complete w2) - 403_1 complete w1, missing w2 
410.2 (missing w1, complete w2) - 410_1 complete w1, missing w2 
434.1 (complete w1, complete w2) - 434_2 with no w2 or w2

506.1 (missing w1, complete w2)
506.2 (missing w1, complete w2)

613.2 (missing w1, complete w2) - 613 complete w1, missing w2 
812.2 (missing w1, complete w2) - 812_1 complete w1, missing w2 

905.1 (complete w1, complete w2)
905.2 (missing w1, complete w2)

914.2 (missing w1, complete w2) - 914 complete w1, missing w2 
939.2 (missing w1, complete w2) - 939 complete w1, missing w2 
1015.2 (missing w1, complete w2) - 1015 complete w1, missing w2 
1027.2 (missing w1, complete w2) - 1027_1 complete w1, missing w2 
1041.2 (missing w1, complete w2) - 1041 complete w1, missing w2 (orange but empty, prob assessor input) 


--------------------------------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------------------------------

# renaming variables (cohort 2)
```{r}

# d1_c2 <- d_c2 %>% 
#   rename(c("demo014_1_5" = "demo014_1_77")) %>% # to match coh_1
#   rename(c("demo019_1_10" = "demo019_1_77")) %>% # to match coh_1
```

# descriptive stats function
```{r include=FALSE}

# function that takes a df and returns a df with only the numeric columns
only_numeric <- function(df) {
  select_numeric <- dplyr::select_if(df, is.numeric)
  return(select_numeric)
}

# a list with the functions we want
smry <- list(n = function(x) length(x),
             n_valid = function(x) sum(!is.na(x)),
             na = function(x) sum(is.na(x)),
             M = function(x) round(mean(x, na.rm = TRUE), 2), 
             SD = function(x) round(sd(x, na.rm = TRUE), 2),
             Min = function(x) round(min(x, na.rm = TRUE), 2),
             Max = function(x) round(max(x, na.rm = TRUE), 2))

# wrapping solution in a function
descriptives <- function(df) {
  select_numeric <- only_numeric(df)
  mean_sd <- map_df(select_numeric, function(col) map_df(smry, ~.x(col)),
       .id = "column")
  return(mean_sd)
}
```