---
title: "initial cleaning"
output: html_document
date: "2023-11-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)

theme_set(theme_minimal())
```

# Loading the dataset (cohort 1 - wave 1)
```{r}
d_raw_w1 <- import(here("nopublish", "Cuidadores Medición 1 Cohorte 1.sav"), setclass = "tbl_df")
```

## WAVE 1

#initial cleaning (wave 1)
```{r}

# d_raw_w1 <- d_raw_c1[-c(1:5), ] # to remove rows of data

d_w1 <- d_raw_w1 %>% 
  select(-4:-6) %>% # selecting out columns with non-relevant data
  clean_names() %>% # function that formats variables' names
  arrange(record_id) # ordering participants ids in descending order
```
# Checking duplicated ids
```{r}
data.frame(table(d_w1$record_id))
```

Duplicate ids:
1038 (confirmed in redcap: one caregiver completed survey for two children boy and girl)
in wave 2 participant only completed survey for 1 child (the boy)

# before recoding N/A
```{r}
view_df(d_w1)
```

I identified the following variables to convert to N/A:
77 = No aplica
98 = no estoy sugura
99 = prefer not to respond

# recoding missing variables as N/A
```{r include=FALSE}

# vector with missing values in dataset
missing_vals <- c(77, 98, 99)

# function that returns true if values in vector are equal to missing_vals. The function takes a vector x, and specified values of missing data
recode_missing <- function(x, missing_vals = c(77, 98, 99)) {
  test <- x %in% missing_vals
  ifelse(test, NA, x)
}

# function that recodes missing values to NA. The function takes a dataframe with variables with missing data, and specified values of missing data
recode_missing_df <- function(df, missing_vals = c(77, 98, 99)) {
  modify(df, ~recode_missing(.x, missing_vals)) # here uses the function created above
}

test <- recode_missing_df(d_w1) # the function strips out variable labels
```

# after recoding N/A
```{r}
view_df(test)
```


## WAVE 2


# Loading the dataset (cohort 1 - wave 2)
```{r}
d_raw_w2 <- import(here("nopublish", "Cuidadores Medición 2 Cohorte 1.sav"), setclass = "tbl_df")
```

#initial cleaning (wave 2)
```{r}

# d_raw_w1 <- d_raw_c1[-c(1:5), ] # to remove rows of data

d_w2 <- d_raw_w2 %>% 
  select(-4:-6) %>% # selecting out columns with non-relevant data
  clean_names() %>% # function that formats variables' names
  arrange(record_id) # ordering participants ids in descending order
```

# Checking duplicated ids
```{r}
data.frame(table(d_w2$record_id))
```
Duplicate ids:
no duplicated id, but oddly named ones:

202.1 (complete w1, complete w2)
202.2 (missing w1, complete w2)

228.2 (missing w1, complete w2) - 228_1 complete w1, missing w2 
403.2 (missing w1, complete w2) - 403_1 complete w1, missing w2 
410.2 (missing w1, complete w2) - 410_1 complete w1, missing w2 
434.1 (complete w1, complete w2) - 434_2 with no w2 or w2

506.1 (missing w1, complete w2)
506.2 (missing w1, complete w2)

613.2 (missing w1, complete w2) - 613 complete w1, missing w2 
812.2 (missing w1, complete w2) - 812_1 complete w1, missing w2 

905.1 (complete w1, complete w2)
905.2 (missing w1, complete w2)

914.2 (missing w1, complete w2) - 914 complete w1, missing w2 
939.2 (missing w1, complete w2) - 939 complete w1, missing w2 
1015.2 (missing w1, complete w2) - 1015 complete w1, missing w2 
1027.2 (missing w1, complete w2) - 1027_1 complete w1, missing w2 
1041.2 (missing w1, complete w2) - 1041 complete w1, missing w2 (orange but empty, prob assessor input) 




--------------------------------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------------------------------

# renaming variables (cohort 2)
```{r}

# d1_c2 <- d_c2 %>% 
#   rename(c("demo014_1_5" = "demo014_1_77")) %>% # to match coh_1
#   rename(c("demo019_1_10" = "demo019_1_77")) %>% # to match coh_1
```

# descriptive stats function
```{r include=FALSE}

# function that takes a df and returns a df with only the numeric columns
only_numeric <- function(df) {
  select_numeric <- dplyr::select_if(df, is.numeric)
  return(select_numeric)
}

# a list with the functions we want
smry <- list(n = function(x) length(x),
             n_valid = function(x) sum(!is.na(x)),
             na = function(x) sum(is.na(x)),
             M = function(x) round(mean(x, na.rm = TRUE), 2), 
             SD = function(x) round(sd(x, na.rm = TRUE), 2),
             Min = function(x) round(min(x, na.rm = TRUE), 2),
             Max = function(x) round(max(x, na.rm = TRUE), 2))

# wrapping solution in a function
descriptives <- function(df) {
  select_numeric <- only_numeric(df)
  mean_sd <- map_df(select_numeric, function(col) map_df(smry, ~.x(col)),
       .id = "column")
  return(mean_sd)
}
```