---
title: "MdM parent scales"
author: "Alejandra Garcia Isaza"
date: "2025-06-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)

theme_set(theme_minimal())
```


### Loading the dataset (cohort 1 and 2 - wave 1)
```{r}
d <- import(here("data", "caregivers_co1_co2_w1.sav"), setclass = "tbl_df")

view_df(d) # Not showing each value label separate. Values appear as range, prob due to haven stripping labels when I exported. 
```


### recoding missing variables as N/A
```{r include=FALSE}

# vector with missing values in dataset
missing_vals <- c(77, 99)

# function that returns true if values in vector are equal to missing_vals. The function takes a vector x, and specified values of missing data
recode_missing <- function(x, missing_vals = c(77, 99)) {
  test <- x %in% missing_vals
  ifelse(test, NA, x)
}

# function that recodes missing values to NA. The function takes a dataframe with variables with missing data, and specified values of missing data
recode_missing_df <- function(df, missing_vals = c(77, 99)) {
  modify(df, ~recode_missing(.x, missing_vals)) # here uses the function created above
}

d1 <- recode_missing_df(d) # the function strips out variable labels
```



female caregivers = 1057
male caregivers = 163
99 = 4 
NA = 1 
total = 1225

intro001_1:
1, Soy su madre = 920
2, Soy su padre = 131
3, Otra relación = 169

demo001_1:
1, Femenino
2, Masculino
99, Prefiero no responder	

Sleuthing:
99 = co1: 835 (tia), co2: 2122 (dad), 2542 (?), 2824(mom) = 2 female, 1 male, 1 ?
NA = co1: 438 (mom) = 1 female

Conclusions of my sleuthing: 
female caregivers = 1060
male caregivers = 164
1 row with 99, id 2542 (?)
total = 1225

```{r}
par_w1
view_df(par_w1)

par_w1 %>%
  	count(demo001_1)

# to identify who are the 99
par_w1 %>%
  	filter(demo001_1 == 99)

# to identify who is the na
par_w1 %>% filter(is.na(demo001_1))

par_w1 %>% filter(record_id == "835") 
par_w1 %>% filter(record_id == "2122") 
par_w1 %>% filter(record_id == "2542") # lots of prefer not to respond
par_w1 %>% filter(record_id == "2824") 
par_w1 %>% filter(record_id == "438") # lots of prefer not to respond

```


### Data prep: reverse scoring negatively worded items

SDQ values = 0, 1, 2
PAFAS values = 0, 1, 2, 3
```{r}
d2 <- d1 %>%
  mutate(sdq007_1 = likert_reverse(sdq007_1, top = 2, bottom = 0),
         sdq021_1 = likert_reverse(sdq021_1, top = 2, bottom = 0),
         sdq025_1 = likert_reverse(sdq025_1, top = 2, bottom = 0),
         sdq011_1 = likert_reverse(sdq011_1, top = 2, bottom = 0),
         sdq014_1 = likert_reverse(sdq014_1, top = 2, bottom = 0),
         pafas003_1 = likert_reverse(pafas003_1, top = 3, bottom = 0),
         pafas011_1 = likert_reverse(pafas011_1, top = 3, bottom = 0),
         pafas019_1 = likert_reverse(pafas019_1, top = 3, bottom = 0),
         pafas020_1 = likert_reverse(pafas020_1, top = 3, bottom = 0),
         couple001_2 = likert_reverse(couple001_2, top = 3, bottom = 0),
         )
```


### checking reverse scoring
```{r}
d1$sdq007_1
d2$sdq007_1
# Conclusion: it worked

d1$pafas019_1
d2$pafas019_1
# Conclusion: it worked
```
