---
title: "students baseline & scales"
author: "Alejandra Garcia Isaza"
date: "2025-06-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rio)
library(here)
library(tidyverse)
library(haven)
library(janitor)
library(knitr)
library(surveytoolbox)
library(sjPlot)
library(kableExtra)
library(psych)
library(sjmisc)

theme_set(theme_minimal())
```


### Loading the dataset (cohort 1 and 2 - wave 1)
```{r}
d_stu <- import(here("data", "students_co1_co2_w1.sav"), setclass = "tbl_df")

view_df(d_stu) # Not showing each value label separate. Values appear as range, prob due to haven stripping labels when I exported. 
```


### recoding missing variables as N/A
```{r include=FALSE}

# vector with missing values in dataset
missing_vals <- c(77, 99)

# function that returns true if values in vector are equal to missing_vals. The function takes a vector x, and specified values of missing data
recode_missing <- function(x, missing_vals = c(77, 99)) {
  test <- x %in% missing_vals
  ifelse(test, NA, x)
}

# function that recodes missing values to NA. The function takes a dataframe with variables with missing data, and specified values of missing data
recode_missing_df <- function(df, missing_vals = c(77, 99)) {
  modify(df, ~recode_missing(.x, missing_vals)) # here uses the function created above
}

d1_stu <- recode_missing_df(d_stu) # the function strips out variable labels
```

### demo variables
```{r}
# age
mean(d1_stu$demo001, na.rm = TRUE)
sd(d1_stu$demo001, na.rm = TRUE)

# gender
# 1 = girl = 602
# 2 = boy = 611
d1_stu %>%
  count(demo002) %>%
  mutate(percent = n / sum(n) * 100)

# grade
d1_stu %>%
  count(demo003) %>%
  mutate(percent = n / sum(n) * 100)

# number of people they live with
mean(d1_stu$demo004, na.rm = TRUE)
sd(d1_stu$demo004, na.rm = TRUE)

d1_stu %>%
  filter(demo004 >= 7) %>%
  nrow()

# have a paid job
d1_stu %>%
  count(demo006) %>%
  mutate(percent = n / sum(n) * 100) # 12% (141)

d1_stu %>%
  filter(demo006input_a >= 10) %>%
  nrow() # 28 work 10 or more hours

# like school
d1_stu %>%
  count(school001) %>%
  mutate(percent = n / sum(n) * 100)

# i think i'll finish high school 1 = not true; 2 = somewhat true; 3 = very true
d1_stu %>%
  count(hope004) %>%
  mutate(percent = n / sum(n) * 100) 

# primary caregiver (student report)
d1_stu %>%
  count(primcaregiver) %>%
  mutate(percent = n / sum(n) * 100) # 73% madre (1); 11.30% abuela (5); 6% padre (3); 4.40% tia (7); 4.04% otra persona (9)
```


### Data prep: reverse scoring negatively worded items
SDQ values = 1, 2, 3
```{r}
d2_stu <- d1_stu %>%
  mutate(sdq007 = likert_reverse(sdq007, top = 3, bottom = 1),
         sdq011 = likert_reverse(sdq011, top = 3, bottom = 1),
         sdq014 = likert_reverse(sdq014, top = 3, bottom = 1),
         sdq021 = likert_reverse(sdq021, top = 3, bottom = 1),
         sdq025 = likert_reverse(sdq025, top = 3, bottom = 1))
```


### checking reverse scoring
```{r}
# d1_stu$sdq007
# d2_stu$sdq007
# Conclusion: it worked
```


### SDQ scales
emotional difficulties: sdq003, sdq008, sdq013, sdq016, sdq024 -> higher, worse emotional problems
conduct difficulties: sdq005, sdq007 (R), sdq012, sdq018, sdq022 -> higher, worse conduct problems
Hyperactivity scale: sdq002, sdq010, sdq015, sdq021 (R), sdq025 (R) -> higher, worse hyperactivity
Peer difficulties scale: sdq006, sdq011 (R), sdq014 (R), sdq019, sdq023 -> higher, worse peer diff
Prosocial scale: sdq001, sdq004, sdq009, sdq017, sdq020 -> higher, better prosociality

```{r}
emo_diff <- d2_stu %>%
  select(sdq003, sdq008, sdq013, sdq016, sdq024) 
alpha(emo_diff) # 0.62

cond_diff <- d2_stu %>%
  select(sdq005, sdq007, sdq012, sdq018, sdq022) 
alpha(cond_diff) # 0.49

hyper_diff <- d2_stu %>%
  select(sdq002, sdq010, sdq015, sdq021, sdq025) 
alpha(hyper_diff) # 0.57

peer_diff <- d2_stu %>%
  select(sdq006, sdq011, sdq014, sdq019, sdq023) 
alpha(peer_diff) # 0.35

prosocial <- d2_stu %>%
  select(sdq001, sdq004, sdq009, sdq017, sdq020) 
alpha(prosocial) # 0.52


tot_diff <- d2_stu %>%
  select(sdq003, sdq008, sdq013, sdq016, sdq024, sdq005, sdq007, sdq012, sdq018, sdq022, sdq002, sdq010, sdq015, sdq021, sdq025, sdq006, sdq011, sdq014, sdq019, sdq023) 
alpha(tot_diff) # 0.73

```

### SDQ alternative scales
```{r}
externalizing <- d2_stu %>%
  select(sdq005, sdq007, sdq012, sdq022, sdq002, sdq010, sdq015) # enoja, no obedece, pelea, roba, inquieto, se mueve, distractible
alpha(externalizing) # 0.64

internalizing <- d2_stu %>%
  select(sdq003, sdq008, sdq013, sdq016, sdq024, sdq019, sdq023) # dolores, preocupado, triste, nervioso, miedos, bullying, prefiere estar con adultos
alpha(internalizing) # 0.64
```
### hope scale
```{r}
hope <- d2_stu %>%
  select(hope001, hope002, hope003, hope004, hope005) 
alpha(hope) # 0.51

# cat(paste(d2_stu$hope006, collapse = "\n"))
```


### CRPBI scale (Children's report of parent behavior Inventory)
original inventory: 30 items, 10 items each subscale: a) support; b) psychological control, c) behavioral control

crpbi001 to crpbi10 items are the complete support subscale
crpbi011 to crpbi17 were part of a surveys used by Erin Rodriguez, but don't really belong to CRPBI. These tap on coercive and punitive parenting.
```{r}

support <- d2_stu %>%
  select(crpbi001, crpbi002, crpbi003, crpbi004, crpbi005, crpbi006, crpbi007, crpbi008, crpbi009, crpbi010) 
alpha(support) # 0.82

neg_coercive <- d2_stu %>%
  select(crpbi011, crpbi012, crpbi013, crpbi014, crpbi015, crpbi016, crpbi017) 
alpha(neg_coercive) # 0.75
```



### code for creating scales
```{r}
student_scales <- d2_stu %>%
  rowwise() %>%
  mutate(sdq_ext = mean(c(sdq005, sdq007, sdq012, sdq022, sdq002, sdq010, sdq015), na.rm = TRUE),
         sdq_int = mean(c(sdq003, sdq008, sdq013, sdq016, sdq024, sdq019, sdq023), na.rm = TRUE),
         support = mean(c(crpbi001, crpbi002, crpbi003, crpbi004, crpbi005, crpbi006, crpbi007, crpbi008, crpbi009, crpbi010), na.rm = TRUE),
         neg_coercive = mean(c(crpbi011, crpbi012, crpbi013, crpbi014, crpbi015, crpbi016, crpbi017), na.rm = TRUE))
```

### Average scores gy gender
1 = girls, 2 = boys
```{r}
# SDQ
student_scales %>%
  group_by(demo002) %>%
  summarise(
    mean_sdq_ext = mean(sdq_ext, na.rm = TRUE),
    sd_sdq_ext = sd(sdq_ext, na.rm = TRUE),
    n = n()
  )

student_scales %>%
  group_by(demo002) %>%
  summarise(
    mean_sdq_int = mean(sdq_int, na.rm = TRUE),
    sd_sdq_int = sd(sdq_int, na.rm = TRUE),
    n = n()
  )

# Support & coercion
student_scales %>%
  group_by(demo002) %>%
  summarise(
    mean_support = mean(support, na.rm = TRUE),
    sd_support = sd(support, na.rm = TRUE),
    n = n()
  )

student_scales %>%
  group_by(demo002) %>%
  summarise(
    mean_neg_coercive = mean(neg_coercive, na.rm = TRUE),
    sd_neg_coercive = sd(neg_coercive, na.rm = TRUE),
    n = n()
  )

```
































### Figures
```{r}
ggplot(student_scales, aes(x = sdq_ext)) + 
  geom_histogram(bins = 30, fill = "cornflowerblue", color = "black", alpha = 0.7) + 
  facet_wrap(~ demo002, scales = "free") +  # One plot per variable
  theme_minimal() + 
  labs(title = "externalizing behavior")
```


```{r}
custom_labels <- c("1" = "Girls", "2" = "Boys")

ggplot(student_scales, aes(x = sdq_ext)) + 
  geom_histogram(bins = 20, fill = "cornflowerblue", color = "black", alpha = 0.7) + 
  facet_wrap(~ demo002, scales = "free", 
             labeller = labeller(demo002 = custom_labels)) +
  theme_minimal() + 
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    strip.text = element_text(size = 12, face = "italic")
  ) +
  labs(
    title = "Externalizing Behavior by Gender"
  )

###
custom_labels <- c("1" = "Girls", "2" = "Boys")

ggplot(student_scales, aes(x = sdq_int)) + 
  geom_histogram(bins = 20, fill = "cornflowerblue", color = "black", alpha = 0.7) + 
  facet_wrap(~ demo002, scales = "free", 
             labeller = labeller(demo002 = custom_labels)) +
  theme_minimal() + 
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    strip.text = element_text(size = 12, face = "italic")
  ) +
  labs(
    title = "Internalizing Behavior by Gender"
  )
```


```{r}
# Step 1: Recode values with labels
figs_stu <- d1_stu %>%
  mutate(demo006_label = case_when(
    demo006 == 0 ~ "No",
    demo006	 == 1 ~ "Yes",
    TRUE ~ "Missing/Other"
  ))

# Step 2: Count and calculate percentages
demo_counts <- figs_stu %>%
  count(demo006_label) %>%
  mutate(percent = round(100 * n / sum(n), 1),
         label = paste0(demo006_label, " (", percent, "%)"))

# Step 3: Create pie chart with percentages
ggplot(demo_counts, aes(x = "", y = n, fill = demo006_label)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y") +
  geom_text(aes(label = paste0(percent, "%")),
            position = position_stack(vjust = 0.5),
            color = "white", size = 4.5) +
  theme_void() +
  labs(title = "Percentage of Children with a Paying Job?",
       fill = "Job?") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        legend.title = element_text(face = "bold"))
```




































########### EFA for funsies ###########

## The whole scale
```{r}
sdq <- d2_stu %>%
  select(sdq003, sdq008, sdq013, sdq016, sdq024, sdq005, sdq007, sdq012, sdq018, sdq022, sdq002, sdq010, sdq015, sdq021, sdq025, sdq006, sdq011, sdq014, sdq019, sdq023, sdq001, sdq004, sdq009, sdq017, sdq020)
```

## kmo Kaiser-Meyer-Olkin factor sampling adequacy (above 0.5, closer 1 is best)
```{r}
KMO(sdq) # 0.81 --> ok
```

```{r}
poly_corr <- polychoric(sdq)$rho
cortest.bartlett(poly_corr, n = nrow(sdq))
```


## Poly corr matrix
```{r}
poly <- polychoric(sdq)
cor.plot(poly$rho, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```

## scree plot
```{r}
scree(sdq, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
# Scree plot shows 3 factor solution, but 1 eigenvalue would be less than 1.0
```

## EFA using ULS 5 factor
```{r}
factor_test_uls <- fa(sdq, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 5) 
factor_test_uls
```

## Omega 
```{r}
omega(sdq, nfactors = 5, flip = T, plot = T) # alpha = 0.76
```
  
  
### Just total difficulties scale

```{r}
sdq <- d2_stu %>%
select(sdq003, sdq008, sdq013, sdq016, sdq024, sdq005, sdq007, sdq012, sdq018, sdq022, sdq002, sdq010, sdq015, sdq021, sdq025, sdq006, sdq011, sdq014, sdq019, sdq023) 
```


## kmo Kaiser-Meyer-Olkin factor sampling adequacy (above 0.5, closer 1 is best)
```{r}
KMO(sdq) # 0.81 --> ok
```

```{r}
poly_corr <- polychoric(sdq)$rho
cortest.bartlett(poly_corr, n = nrow(sdq))
```


## Poly corr matrix
```{r}
poly <- polychoric(sdq)
cor.plot(poly$rho, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```

## scree plot
```{r}
scree(sdq, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
# Scree plot shows 3 factor solution, but 1 eigenvalue would be less than 1.0
```

## EFA using ULS 3 factor
```{r}
factor_test_uls <- fa(sdq, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 4) 
factor_test_uls
```

## Omega 
```{r}
omega(sdq, nfactors = 5, flip = T, plot = T) # alpha = 0.76
```
  
## Just the 

externalizing <- d2_stu %>%
  select(sdq005, sdq007, sdq012, sdq022, sdq002, sdq010, sdq015) # enoja, no obedece, pelea, roba, inquieto, se mueve, distractible
alpha(externalizing) # 0.64

internalizing <- d2_stu %>%
  select(sdq003, sdq008, sdq013, sdq016, sdq024, sdq019, sdq023) # dolores, preocupado, triste, nervioso, miedos, bullying, prefiere estar con adultos
alpha(internalizing) 


```{r}
sdq <- d2_stu %>%
  select(sdq003, sdq008, sdq013, sdq016, sdq024, sdq005, sdq012, sdq022, sdq002, sdq010, sdq015, sdq011, sdq014)
```

## kmo Kaiser-Meyer-Olkin factor sampling adequacy (above 0.5, closer 1 is best)
```{r}
KMO(sdq) # 0.75 --> ok
```

```{r}
poly_corr <- polychoric(sdq)$rho
cortest.bartlett(poly_corr, n = nrow(sdq))
```
## Poly corr matrix
```{r}
poly <- polychoric(sdq)
cor.plot(poly$rho, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```

## scree plot
```{r}
scree(sdq, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
# Scree plot shows 3 factor solution, but 1 eigenvalue would be less than 1.0
```

## EFA using ULS 3 factor
```{r}
factor_test_uls <- fa(sdq, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 3) 
factor_test_uls
```


```{r}
externalizing <- d2_stu %>%
  select(sdq005, sdq012, sdq022, sdq002, sdq010, sdq015) 
alpha(externalizing) # 0.62

internalizing <- d2_stu %>%
  select(sdq003, sdq008, sdq013, sdq016, sdq024) 
alpha(internalizing) # 0.62
```

### attempt 2

## The scale to factor analyze
```{r}
sdq <- d2_stu %>%
  select(sdq003, sdq008, sdq013, sdq016, sdq024, sdq005, sdq007, sdq012, sdq022, sdq002, sdq010, sdq015,  sdq011, sdq014, sdq019, sdq023)
```

## kmo Kaiser-Meyer-Olkin factor sampling adequacy (above 0.5, closer 1 is best)
```{r}
KMO(sdq) # 0.81 --> ok
```

```{r}
poly_corr <- polychoric(sdq)$rho
cortest.bartlett(poly_corr, n = nrow(sdq))
```
## Poly corr matrix
```{r}
poly <- polychoric(sdq)
cor.plot(poly$rho, numbers = T, upper = F, main = "Polychoric", show.legend = F)
```

## scree plot
```{r}
scree(sdq, factors = TRUE, pc = FALSE, main = "Scree plot", hline = NULL, add = FALSE)
# Scree plot shows 3 factor solution, but 1 eigenvalue would be less than 1.0
```

## EFA using ULS 3 factor
```{r}
factor_test_uls <- fa(sdq, n.obs = 95, rotate = "oblimin", fm = "uls", cor = "poly", nfactors = 3) 
factor_test_uls
```

removed:
18 acusado de mentir
21 piensa antes
25 termina lo q
6 prefiere solo

just 2 items in one factor
sdq011 al menos 1 amigo
sdq014 cae bien